Aligning marine species range data to better serve science and conservation

### 2016-08-28 clip_fao/fao_clip_pac.Rmd
Omitted OHI analysis from intro, methods, and res/disc/concl; added results/discussion for AM FAO clipping

@@

clipped_spp <- am_spp_central_lims %>%
  # bind_rows(am_spp_north_lims) %>%
  # filter(long %in% c(77, 185)) %>%
  filter(long == 185) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(spp_quad_list, by = 'am_sid')

@@

rast_list <- vector('list', length = nrow(spp_examples_e))

for (i in 1:nrow(spp_examples_e)) {  # i <- 1

  spp <- spp_examples_e[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  r1 <- r_am_spp %>%
    crop(extent(c(-180, 0, -50, 50)))
  extent(r1) <- c(180, 360, -50, 50)

  r2 <- r_am_spp %>%
    crop(extent(c(0, 180, -50, 50)))

  r3 <- merge(r2, r1)

  new_ext <- extent(c(50, 240, -50, 50))
  r_am_crop <- crop(r3, new_ext)

  rast_list[i] <- r_am_crop
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_crop, spp$sciname)
}

mapstack_e <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)
plot(mapstack_e)
values(mapstack_e)[values(mapstack_e) == 0] <- NA
mapstack_e <- crop(mapstack_e, extent(c(85, 220, -50, 50)))

stack_plot_e <- assemble_map(mapstack_e, 'n_spp') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot_e)

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim') %>%
  sample_n(30)

rast_list <- vector('list', length = nrow(spp_examples_w))

for (i in 1:nrow(spp_examples_w)) {  # i <- 1

  spp <- spp_examples_w[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  r1 <- r_am_spp %>%
    crop(extent(c(-180, 0, -50, 50)))
  extent(r1) <- c(180, 360, -50, 50)

  r2 <- r_am_spp %>%
    crop(extent(c(0, 180, -50, 50)))

  r3 <- merge(r2, r1)

  new_ext <- extent(c(70, 220, -50, 50))
  r_am_crop <- crop(r3, new_ext)

  rast_list[i] <- r_am_crop
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_crop, spp$sciname)
}

mapstack_w <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)
plot(mapstack_w)
values(mapstack_w)[values(mapstack_w) == 0] <- NA
mapstack_w <- crop(mapstack_w, extent(c(85, 220, -50, 50)))

stack_plot_w <- assemble_map(mapstack_w, 'n_spp') +
  tm_shape(fao_polys) +
  tm_borders()

print(stack_plot_w)

>>>>

am_spp_list <- read_csv(file.path(dir_git, 'data/spp_all.csv')) %>%
  filter(!is.na(am_sid)) %>%
  arrange(reviewed) %>%
  filter(!duplicated(am_sid)) %>%
  select(am_sid, sciname, reviewed) %>%
  distinct()

@@

clipped_spp <- am_spp_central_lims %>%
  # bind_rows(am_spp_north_lims) %>%
  # filter(long %in% c(77, 185)) %>%
  filter(long == 185) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(am_spp_list, by = 'am_sid')

@@

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim')

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum
spp_cells_clipped <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_w$am_sid) %>%
  group_by(loiczid) %>%
  summarize(n_spp = n())

rast_spp_w <- merge(rast_spp, rast_spp_shift)

values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
# values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
rast_spp_w <- crop(rast_spp_w, extent(c(85, 220, -59, 50)))

rast_spp_map_w <- 
  tm_shape(rast_spp_w) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 200, 300, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Count',
              alpha = 1) +
  tm_shape(World) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_CODE_1',
            col = 'orangered4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .35) +
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            legend.text.size = .35,
            legend.title.size = .45,
            # title.position = 'TOP', 
            legend.outside = TRUE,
            # legend.position = c('right', 'bottom'),
            # legend.bg.color = 'white',
            # legend.bg.alpha = .5,
            attr.outside = TRUE)

print(rast_spp_map_w)

save_tmap(rast_spp_map_w, file.path(dir_fig, 'fig4_am_edge_w.png'), 
          width = 9, height = 6, units = 'cm', dpi = 600)

### 2016-08-28 ms_3_res_disc_concl.Rmd
updated aquamaps edge map
Omitted OHI analysis from intro, methods, and res/disc/concl; added results/discussion for AM FAO clipping

[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges is likely due to experts taking a conservative and precautionary approach.  But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.  Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.  While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.

## Georegion constraint exploration

@@

[results]

## Case Studies

@@

### Case Study: The Ocean Health Index

@@

### Case Study: MPA Gap Analysis

>>>>

[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges may be due to experts taking a precautionary approach.  But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.  Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.  While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.

## Georegional constraint exploration

@@

[results] We identified 4,168 species whose equatorial ranges encounter a range limit at 175° W.  This longitude created a western limit for 512 species, and eastern limit for 3,656.  Plotting the entire ranges of the 512 west-limited species results in a very clear exclusion zone that matches perfectly with FAO region 71; while we did not focus on the boundary for region 57, a similar effect is readily apparent.

[discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, enabling rapid modeling of thousands of species ranges.  However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd species range predictions where otherwise suitable habitat encounters border between regions.  While such an odd boundary would likely be obvious when inspecting ranges of individual species, the distinction is likely to be obscured when many species ranges are aggregated as is typical for biodiversity or conservation studies.  Applying an additional spatial constraint based upon ecological or environmental boundaries, such as Marine Ecoregions of the World [@REF???], may help boost confidence in these predicted ranges; expert review, though time-consuming, is certain to improve confidence in these predicted ranges.

<!-- ## Case Studies

@@

-->

<!--
### Case Study: The Ocean Health Index

@@

-->

## Case Study: MPA Gap Analysis
